Natural sources of vagueness and their implications

### 2017-01-23 book_chapter/paper.tex
Classical problem statement.

\documentclass{article}
\usepackage{todonotes}
\usepackage{a4wide}
\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false]{biblatex}
\bibliography{paper}

@@

\section{Vaguenes}
\label{sec:vagueness}

>>>>

\documentclass[a4paper]{article}

\usepackage[natbib=true,style=authoryear-comp,backend=biber,doi=false,url=false]{biblatex}
\bibliography{paper}

\usepackage{todonotes}
\usepackage{a4wide}
\usepackage{amssymb}
\usepackage{syllogism}

@@

\section{Vagueness}
\label{sec:vagueness}

The classical philosophical problem of vagueness is most starkly embodied by the Sorites paradox.
The original formulation is attributed to Eubulides, an ancient Megarian philosopher~\citep{sorensen_sorites_2009}, and uses the example of a heap of sand: if one million grains of sand piled up together clearly form a heap, and removing one grain of sand can never make a heap into a non-heap, then by repeated applications of this reasoning we are lead to acknowledge that one single grain of sand also forms a heap.
This paradox can be stated in a very general way.
We start with an $x$ to which a certain predicate $P$ clearly applies.
We observe that a certain transformation $f$ has no impact on whether something is or is not $P$.
Thus we are lead to acknowledge that the predicate $P$ also applies to $y = f(x)$.
Moreover, we can apply the same reasoning subsequently to $f(x)$ and repeat that as many times as the transformation is still applicable.
The paradox arises when $n$ repeated applications of the transformation $f$ lead us from $x$ to $z = f^n(x)$, where $z$ is something for which $P$ clearly does not apply.
However, the original premisses seem to force us to accept that $P$ does apply to $x$.
The paradox with the heap can thus be seen as an instance of this general formulation, if $x$ is the group of grains of sand, the transformation is removing a grain, and $z$ the state of one or zero grains.
It can also be reversed: it seems reasonable to accept that piling up a grain of sand to something that is not a heap will not turn it into a heap of sand, thus 
by repeated applications of that transformation one can pile up as many grains of sand as one likes and will never obtain a heap.

Predicates for which one can find a suitable instance of the general formulation of the Sorites paradox are called \emph{vague}.
Paradigmatic examples besides `heap' include `tall', `red', `bald', `tadpole', and `child'~\citep{Keefe1997}.
An important intuition regarding these predicates that enables the Sorites paradox to work is that, for a certain level of granularity, there is no clear boundary demarcating the cases where the predicate applies from the cases where it doesn't, but there are nevertheless examples of both cases.
Intuitively, a person with a height of 2 meters is tall, and one with a height of 1.5 meters is not, but there is no value of height in between where one would draw the line; saying, for example, that a person with a height of 1.80 meters is tall, but a person measuring 1.79 meters is not, just seems preposterous.
Neither drawing a sharp boundary seems reasonable, nor do the consequences of not doing so, and in there lies the truely paradoxical issue of vagueness.
How widespread is the problem?
Mereological nihilists argue that instances of the Sorites can be designed for any material object that can be decomposed into small enough parts.
If we subscribe to the scientific picture of material objects as composed of molecules and atoms, this applies to tables and chairs, cats and mats, and any other ordinary thing~\citep{Unger1979}.
Any heap of molecules that we have a name for is potentially at the mercy of being hypotheticall chipped away into nothingness while still keeping its name.
The problem thus seems to be a serious one.
In face of such a powerful paradox like the Sorites, something has got to give.
